\documentclass[12pt]{article}

% Packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

% Title
\title{CS312 Project 1 Report}
\author{Brock Gregersen}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\newpage


\section{Algorithm Analysis}

\subsection{Time and Space Complexity}
In this section, explain the time and space complexity of your algorithm by breaking down each part of your code. 

\subsubsection{Prime Number Generation}
Discuss the complexity of generating prime numbers using a method like trial division or Sieve of Eratosthenes.

\subsubsection{Fermat Primality Test}
Explain the time complexity of the Fermat test, including how many iterations you run and the cost per iteration.

\subsubsection{Miller-Rabin Primality Test}
Break down the Miller-Rabin test's complexity in terms of both time and space.

\subsubsection{Extended Euclidean Algorithm}
Include the complexity of the Extended Euclidean Algorithm for finding the multiplicative inverse.

\paragraph{Summary of Complexity}
Sum up the time and space complexities of the entire algorithm.

\section{Correctness Probability of Primality Tests}
Discuss the equation you used to compute the probability of correctness for both the Fermat and Miller-Rabin implementations.

\subsection{Fermat Primality Test}
Provide the formula for the probability of correctness and explain why it works.

\subsection{Miller-Rabin Primality Test}
Provide the formula for the probability of correctness and explain the significance of each term.


\end{document}